```{r Initialization, include=FALSE}
options(width = 200)
load.project()
```
# What If Everyone Lived in the Big City? Round 1

On a recent roadtrip I was driving through rural Alabama and wondered what the country would look like if we emptied out the rural areas and all moved to cities. I'm working on learning R and though this might be a useful exercise to practice with.

## US Cities Data

I'm starting with a small matrix of US cities that will be the basis for a pairwise distance matrix. This is a default data set privided by the maps library. This matrix has `r length(loc[,1])` cities and the population sums to `r format(sum(loc[,1]), scientific=FALSE, big.mark=",")`. Obviously this list does not have a full tally of the residents in the USA!

```{r City Data}
head(loc)
dm <- rdist.earth(loc[,2:3], loc[,2:3], miles=TRUE)

# Setup names and zero out small values
rownames(dm) <- name
colnames(dm) <- name
dm[dm[] < 0.1] <- 0

# A sample of the data
dm[1:5,1:5]
```

## How do we move people?

The algoritm is that the following function will select the smallest remaining city and move its population to the closest city.

```{r Move Smallest to Closest Function, child="lib/movesmallestToClosest.rmd"}
```

Here's an example:

```{r Example Move}
temppop <- pop
temppop <- moveSmallestToClosest(temppop, dm, verbose=TRUE)
```

## Move everyone!

Now lets start moving people until every city has over 1 million people.

```{r Move Everyone}
repeat {
  pop <- moveSmallestToClosest(pop, dm)
  
  # Get the latest pass
  pop.pass <- pop[,ncol(pop)]
  
  # Inspect just the cities that still have population
  pop.pass <- pop.pass[pop.pass > 0]
  smallcities <- length(pop.pass[pop.pass < 1e6])
  #printf("Small cities remaining: %d", smallcities)
                      
  if (smallcities == 0) {
    break
  }
}
```

This completes in `r ncol(pop)-1` passes and we end up with `r nrow(pop[pop[,ncol(pop)] > 0,])` cities:

```{r Table}
pop[pop[,ncol(pop)] > 0, c(1, ncol(pop))]
```

## Map

This maps shows each of the original cities as a grey dot. The new cities with over a million residents each are shown as a blue circle:

```{r Map}
map('state')
points(loc[,2], loc[,3], col='grey', pch=".", cex=5)
points(loc[pop[,ncol(pop)] > 0,2], loc[pop[,ncol(pop)] > 0,3], col='blue')
```

# References

* http://stackoverflow.com/questions/8509329/finding-nearest-neighbor-between-2-sets-of-dated-points